05_DESeq_manipulation

### 2019-05-20 analysis/05_DESeq2_manipulate.Rmd
half way there

```{r dds}
## subset for test 
#m.colData <- m.colData %>%
#  filter(sextissue == "male_hypothalamus") %>%
# droplevels()

levels(m.colData$sextissue)

for (eachgroup in levels(m.colData$sextissue)){

  print(eachgroup)
  dds <- subsetDESeq(eachgroup)

  dfname <- paste("dds", eachgroup, sep = "")
  assign(paste("dds", eachgroup, sep = ""), dds)

  return(dfname)
}

@@

```{r totalDEGs}
#mydds <- list(ddsmale_hypothalamus) # for testing only
mydds <- list(ddsfemale_hypothalamus, ddsfemale_pituitary, ddsfemale_gondad,
              ddsmale_hypothalamus, ddsmale_pituitary, ddsmale_gondad)

#create list of groups
a <- levels(colData$treatment)
b <- levels(colData$treatment)

  # comapre all contrasts, save to datafrmes
  totalDEGS=data.frame()
  for (i in a){
    for (j in b){
      if (i != j) {
        k <- paste(i,j, sep = ".") #assigns usique rownames
        print(k)
        totalDEGS[k,1]<-i               
        totalDEGS[k,2]<-j
        totalDEGS[k,3]<- numDEGs(i,j) #caluculates number of DEGs
      }
    }
        b <- b[-1]  # drop 1st element of second string to not recalculate DEGs
  }

  totalDEGSname <- paste("totalDEGS", eachgroup, sep = "")
  assign(paste("totalDEGS", eachgroup, sep = ""), totalDEGS)
  return(totalDEGSname)
}

a <- plottotalDEGs(totalDEGSfemale_hypothalamus, "female hypothalamus")
b <- plottotalDEGs(totalDEGSfefemale_pituitary, "female pituitary")
c <- plottotalDEGs(totalDEGSmale_gonad, "female gonad")
d <- plottotalDEGs(totalDEGSmale_hypothalamus, "male hypothalamus")
e <- plottotalDEGs(totalDEGSmale_pituitary, "male pituitary")
f <- plottotalDEGs(totalDEGSmale_gonad, "male gonad")

@@

@@

  pca1 <- ggplot(pcadata, aes(treatment, PC1,color = treatment)) + 
    geom_boxplot() +
    ylab(paste0("PC1: ", percentVar[1],"% variance")) +
    xlab(NULL) +
    theme_cowplot(font_size = 8, line_size = 0.25) +
    labs(subtitle = eachgroup) +
    theme(legend.position = "none")

  pca2 <- ggplot(pcadata, aes(treatment, PC2,color = treatment)) + 
    geom_boxplot() +
    ylab(paste0("PC2: ", percentVar[2],"% variance")) +
    xlab(NULL) +
    theme_cowplot(font_size = 8, line_size = 0.25) +
    labs(subtitle = eachgroup) +
    theme(legend.position = "none")

@@

```{r pheatmap}

@@

  a <- levels(colData$treatment)
  b <- levels(colData$treatment)

@@

  sigDEGs <- as.matrix(sigDEGs) 
  pheatmap(sigDEGs, show_rownames = F, show_colnames = F,
         color = viridis(30),
         breaks=myBreaks,
         annotation_col=anndf,
         main = eachgroup)

  pheatmap(sigDEGs, kmeans_k = 5,
         show_rownames = F, show_colnames = F,
         color = viridis(30),
         breaks=myBreaks,
         annotation_col=anndf,
         main = eachgroup)
}

@@

candidates <- full_join(geneinfo, DEGs)
drop.cols <-colnames(candidates[,grep("padj|pval|pmin", colnames(candidates))])
candidates <- candidates %>% dplyr::select(-one_of(drop.cols))
candidates <- candidates %>%
  filter(Name %in% c( #"JUN", "JUND", "EGR",  "AVP", "AVPR1A", "AVPR1B", "AVPR2", "OXT",  
                     "AR", "CYP19A1", "ESR1", "ESR2", "FSHR",
                     "GHRL", "GAL", "NPVF", "GNRH1", "LHCGR",
                     "PGR", "PRL", "PRLR", "VIP", "VIPR1")) 
row.names(candidates) <- candidates$Name
candidates <- candidates %>% select(-row.names, -rownames, -Name, -geneid)
candidates <- candidates %>% drop_na()
candidates <- as.data.frame(t(candidates))
candidates$RNAseqID <- rownames(candidates)
candidates <- candidates %>% gather(gene, value, -RNAseqID)  %>% 
  filter(RNAseqID != "gene")
candidates$value <- as.numeric(candidates$value)
candidates$V1  <- candidates$RNAseqID
candidatecounts <- left_join(candidates, colData)
candidatecounts$faketime <- as.numeric(candidatecounts$treatment)
candidatecounts$gene <- as.factor(candidatecounts$gene)

p1 <- candidatecounts %>%
  filter(gene %in% c( "AR", "CYP19A1", "ESR1", "ESR2", "FSHR",
                      "GHRL", "GAL", "NPVF", "GNRH1", "LHCGR",
                      "PGR", "PRL", "PRLR", "VIP", "VIPR1")) %>%
  ggplot(aes(x = treatment, y = value, fill = treatment)) +
  geom_boxplot() +
  facet_wrap(~gene, scales = "free") +
  theme_bw(base_size = 8) +
  theme(axis.text.x = element_blank(),
        legend.position = "bottom") +
  labs(subtitle = eachgroup, x = NULL) +
  guides(fill= guide_legend(nrow=1))
p1

>>>>

```{r dds, cache = T}
dds.female_hypothalamus <- subsetDESeq("female_hypothalamus")
dds.female_pituitary <- subsetDESeq("female_pituitary" )
dds.female_gonad <- subsetDESeq("female_gonad" )
dds.male_hypothalamus <- subsetDESeq("male_hypothalamus" )
dds.male_pituitary <- subsetDESeq("male_pituitary"  )
dds.male_gondad <- subsetDESeq("male_gonad")
```

@@

```{r totalDEGs, eval = F}
#create list of groups for deseq contrasts
group1 <- c("m.inc.d3","m.inc.d8", "m.inc.d9", "m.inc.d17", "prolong","extend", "m.n2" ) 
group2 <- group1

a <- plottotalDEGs(dds.female_hypothalamus, "female hypothalamus")
b <- plottotalDEGs(dds.female_pituitary, "female pituitary")
c <- plottotalDEGs(dds.female_gonad, "female gonad")
d <- plottotalDEGs(dds.male_hypothalamus, "male hypothalamus")
e <- plottotalDEGs(dds.male_pituitary, "male pituitary")
f <- plottotalDEGs(dds.male_gondad, "male gonad")

@@

#mydds <- list(ddsmale_hypothalamus) # for testing only
mydds <- list(dds.female_hypothalamus, dds.female_pituitary, dds.female_gonad,
             dds.male_hypothalamus, dds.male_pituitary, dds.male_gondad)

@@

  pca1 <- ggplot(pcadata, aes(treatment, PC1,color = treatment)) + 
    geom_boxplot() +
    ylab(paste0("PC1: ", percentVar[1],"% variance")) +
    xlab(NULL) +
    theme_cowplot(font_size = 8, line_size = 0.25) +
    labs(subtitle = "eachgroup") +
    theme(legend.position = "none")

  pca2 <- ggplot(pcadata, aes(treatment, PC2,color = treatment)) + 
    geom_boxplot() +
    ylab(paste0("PC2: ", percentVar[2],"% variance")) +
    xlab(NULL) +
    theme_cowplot(font_size = 8, line_size = 0.25) +
    labs(subtitle = "eachgroup") +
    theme(legend.position = "none")

@@

```{r pheatmap, eval = F}

@@

  a <- levels(m.colData$treatment)
  b <- a

@@

  sigDEGs <- as.matrix(sigDEGs) 
  pheatmap(sigDEGs, show_rownames = F, show_colnames = F,
         color = viridis(30),
         breaks=myBreaks,
         annotation_col=anndf,
         main = "eachgroup")

  pheatmap(sigDEGs, kmeans_k = 5,
         show_rownames = F, show_colnames = F,
         color = viridis(30),
         breaks=myBreaks,
         annotation_col=anndf,
         main = "eachgroup")
}

@@

candidates <- full_join(geneinfo, DEGs)
drop.cols <-colnames(candidates[,grep("padj|pval|pmin", colnames(candidates))])
candidates <- candidates %>% dplyr::select(-one_of(drop.cols))
candidates <- candidates %>%
  filter(Name %in% c( #"JUN", "JUND", "EGR",  "AVP", "AVPR1A", "AVPR1B", "AVPR2", "OXT",  
                     "AR", "CYP19A1", "ESR1", "ESR2", "FSHR",
                     "GHRL", "GAL", "NPVF", "GNRH1", "LHCGR",
                     "PGR", "PRL", "PRLR", "VIP", "VIPR1")) 
row.names(candidates) <- candidates$Name
candidates <- candidates %>% select(-row.names, -rownames, -Name, -geneid)
candidates <- candidates %>% drop_na()
candidates <- as.data.frame(t(candidates))
candidates$RNAseqID <- rownames(candidates)
candidates <- candidates %>% gather(gene, value, -RNAseqID)  %>% 
  filter(RNAseqID != "gene")
candidates$value <- as.numeric(candidates$value)
candidates$V1  <- candidates$RNAseqID

candidatecounts <- left_join(candidates, m.colData)
candidatecounts$faketime <- as.numeric(candidatecounts$treatment)
candidatecounts$gene <- as.factor(candidatecounts$gene)

p1 <- candidatecounts %>%
  filter(gene %in% c( "AR", "CYP19A1", "ESR1", "ESR2", "FSHR",
                      "GHRL", "GAL", "NPVF", "GNRH1", "LHCGR",
                      "PGR", "PRL", "PRLR", "VIP", "VIPR1")) %>%
  ggplot(aes(x = treatment, y = value, fill = treatment)) +
  geom_boxplot() +
  facet_wrap(~gene, scales = "free") +
  theme_bw(base_size = 8) +
  theme(axis.text.x = element_blank(),
        legend.position = "bottom") +
  labs(x = NULL) +
  guides(fill= guide_legend(nrow=1))

plot(p1)
}
```
